Classical Mechanics Assignment 5

### 2015-10-29 classical/latex/a5.tex
Add Classical HW 5

\title{Classical Mechanics Assignment 4}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One}
\textbf{Consider a rocket accelerating in free space by burning its fuel at a constant rate. Find the velocity at which the momentum of the rocket (neglecting the already spent fuel) is maximized.} \\

We want to find the velocity at which $m\dot{y}$ is maximized. We find when the derivative of this with respect to time is zero, plug in for the values we find and go from there:

\begin{align*}
  \frac{\partial}{\partial t} m\dot{y} &= \dot{m}\dot{y} + m\ddot{y} = 0\\
  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
  \dot{m}\dot{y} &= mg + \dot{m}v_{ex}\\
  \dot{y} &= \frac{mg + \dot{m}v_{ex}}{\dot{m}}\\
\end{align*}

\section{Problem Two}
\textbf{Consider a rocket subject to a linear resistive force $\vec{F} = -b\vec{v}$, but no other external forces. Find the velocity as a function of the remaining mass for a rocket that starts from rest and ejects mass at a constant rate $k = -\dot{m}$ and velocity $v_{ex}$.} \\

\begin{align*}
  m\ddot{y} &= -\frac{b}{m}\dot{y} - \dot{m}v_{ex}\\
\end{align*}

\section{Problem Three}
\textbf{Consider the surface generated by revolving a line connecting two points ($x_1$,$y_1$) and ($x_2$,$y_2$) about an axis coplanar with the two points.  Find the equation of the line connecting the points such that the surface area generated by the revolution is minimum.}

\begin{align*}
  SA &= \int_{y_1}^{y_2} dA = \int_{y_1}^{y_2} 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}dy\\
  f(x, x', y) &= 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}\\
\end{align*}

\begin{align*}
  \frac{\partial f}{\partial x} &- \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} = 0\\
  \frac{\partial f}{\partial x} &= \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}}\\
  \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} &= 0\\
\end{align*}

So...
\begin{align*}
  \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}} = 0\\
\end{align*}

\section{Problem Four}
\textbf{Consider a medium in which the refractive index $n$ is inversely proportional to $r_2$. That is, $n=a/r2$, where $r$ is the distance from the origin. Use Fermat’s principle to find the path of light traveling between two points that lie in the xy plane. Hint: use 2-dimensional polar coordinates. Show that the path found is part of a circle that intersects the origin of the reference frame.}

We use polar coordinates to describe our two points. Our differential is $ds = \sqrt{r^2+r'^2}d\theta$.

\begin{align*}
  T &= \int_{\theta_1}^{\theta_2} \frac{1}{v}ds = \int_{\theta_1}^{\theta_2} \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
  f(r,r',\theta) &= \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
\end{align*}

\begin{align*}
  \frac{\partial f}{\partial r} - \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
  \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
\end{align*}

And so:

\begin{align*}
    \frac{\partial f}{\partial r} = 0\\
\end{align*}

Meaning

\end{document}

>>>>

\title{Classical Mechanics Assignment 5}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One}
\textbf{Find the shape assumed by an infinitely flexible rope of linear density $\rho$ and length L with both sides attached to the ceiling of a room. (Hint, the shape will minimize the gravitational potential energy of the rope).}

\begin{align*}
  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\right]dx\\
  f(y,y',x) &= -gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\\
\end{align*}

We use the $\frac{df}{dx'} = k$ form of the Euler-Lagrange equation:

\begin{align*}
  \frac{-gy\rho x'}{\sqrt{1 + x'^2}} - \frac{\lambda x'}{\sqrt{1+x'^2}} &= k\\
  \frac{-gy\rho x' - \lambda x'}{\sqrt{1 + x'^2}} &= k\\
\end{align*}

\section{Problem Two}
\textbf{Demonstrate that the shortest distance between two points in 3D is a straight line.}

\begin{align*}
  L &= \int_{z_1}^{z_2} \sqrt{1 + x'^2 + y'^2} dz\\
  f(x, x', y, y', z) &= \sqrt{1 + x'^2 + y'^2}\\
\end{align*}

Since there is no x, y or z dependence in $f$, we can use the condensed Euler-Lagrange equation $\frac{df}{dn'} = k$:\\

\begin{align*}
  \frac{df}{dx'} &= K_1\\
  \frac{x'}{\sqrt{1 + x'^2 + y'^2}} &= K_1\\
  x'^2 &= K_1^2\left(1 + x'^2 + y'^2\right)\\
  x'^2 &= \frac{K_1^2\left(1 + y'^2\right)}{1-K_1}\\
  y'^2 &= \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}
  \hspace*{1in} \mbox{Solutions for x' and y' are symmetric}\\
  x'^2 &= \frac{K_1^2\left(1 + \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}\right)}{1-K_1}\\
  x'^2\left(1-K_1\right) &= K_1^2\left(1 + \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}\right)\\
  x'^2\left(1-K_1\right) &= K_1^2 + \frac{K_1^2K_2^2\left(1 + x'^2\right)}{1-K_2}\\
  x'^2\left(1-K_1\right) &= K_1^2 + \frac{K_1^2K_2^2}{1-K_2} + \frac{x'^2K_1^2K_2^2}{1-K_2}\\
  x'^2\left(1-K_1 - \frac{K_1^2K_2^2}{1-K_2}\right) &= K_1^2 + \frac{K_1^2K_2^2}{1-K_2}\\
  x' &= \sqrt{\frac{K_1^2 + \frac{K_1^2K_2^2}{1-K_2}}{1-K_1 - \frac{K_1^2K_2^2}{1-K_2}}}\\
\end{align*}

Because $x'$ is constant and $y'^2 = \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}$, y' is also constant. Thus we have the equation of a line!\\

This was a fun easy problem. It was interesting that it was so easy to add arbitrary numbers of variables to the Euler-Lagrange equation. To add a variable, you just add another Euler-Lagrange equation to the system. It is strange that the equations are not coupled at all. It kind of makes sense, since you are just minimizing the path in two dimensions for each equation. You could take the 3D plot and look at its XZ side and minimize that. Then you could minimize the YZ plot. The XZ and YZ plots describe the same curve, but are not related.
\end{document}
